---
title: "p8105_hw6_si2426"
output: github_document
author: "Shodai Inose"
date: "2022-11-30"
---

```{r, include = FALSE}
library(tidyverse)
library(purrr)
library(epiDisplay)
library(glmnet)
library(modelr)
library(mgcv)
```

## Problem 2

#### Loading and Cleaning Data

```{r, warning = FALSE}
homicide_df = read.csv("./data/homicide-data.csv") %>% 
  mutate(city_state = paste(city, state, sep = ", "), 
         solved = ifelse(disposition == "Closed by arrest", 1, 0), victim_age = as.numeric(victim_age), victim_race = fct_relevel(victim_race, "White")) %>%
  filter(city_state != "Dallas, TX", city_state != "Phoenix, AZ",
           city_state != "Kansas City, MO",
           city_state != "Tulsa, AL") %>%
  filter(victim_race ==  "White" | victim_race ==  "Black")
```

The data was wrangled and cleaned by creating a variable for `city_state` and a binary variable `solved` to identify cases that were closed by an arrest. Several cities were filtered out and only victims of White or Black ethnic backgrounds will be abalyzed. 

```{r}
baltimore_homicide = homicide_df %>% 
  filter(city == "Baltimore")

baltimore_logistic = 
  baltimore_homicide %>% 
  glm(solved ~ victim_age + victim_race + victim_sex, data = ., family = binomial()) 

baltimore_logistic_df = 
  baltimore_logistic %>% 
  broom::tidy() %>% 
  mutate(OR = exp(estimate)) 

baltimore_logistic_table = baltimore_logistic_df %>%
  dplyr::select(term, log_OR = estimate, OR, p.value) %>% 
  knitr::kable(digits = 3)

baltimore_logistic_table
```

```{r}
as_tibble(confint(baltimore_logistic)) %>% slice(4) %>% exp()
```

A 95% confidence interval for the adjusted odds ratio for solving homicides in Baltimore, Maryland in which the victim is male (keeping all other variables fixed) is (0.32, 0.56). This indicates that homicides in which the victim is male are less likely to be solved than cases in which the victim is female in Baltimore, Maryland.

```{r, message = FALSE}
cities_logistic = homicide_df %>% 
  group_by(city_state) %>% 
  nest() %>%
  mutate( 
    models = map(data, ~glm(data = .x, solved ~ victim_age + victim_race + victim_sex, family = binomial())),
    result = map(models, broom::tidy)) %>%
  dplyr:: select(-data, -models)  %>% 
  unnest(result) %>%
  mutate(OR = exp(estimate)) %>%
  dplyr:: select(term, log_OR = estimate, std_error = std.error, OR, p.value) %>%
  mutate(lwr = exp(log_OR - 1.96*std_error), upr = exp(log_OR + 1.96*std_error))

cities_logistic =
  cities_logistic[, c("city_state", "term", "log_OR", "std_error", "OR", 
                      "lwr", "upr",
                      "p.value")]
```

```{r}
victim_sex_logistic = cities_logistic %>% 
  filter(term == "victim_sexMale") %>%
  ungroup(city_state) %>%
  mutate(city_state = fct_reorder(city_state, OR))

ggplot(victim_sex_logistic, aes(x = city_state, y = OR)) + 
  geom_point() +
  geom_errorbar(aes(ymin = lwr, ymax = upr)) +
  labs(x = "City, State", y = "Adjusted Odds Ratio") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```

## Problem 3

#### Loading and Cleaning Data
```{r}
birthweight_df = read.csv("./data/birthweight.csv") %>%
  mutate(babysex = as.factor(babysex),
         frace = as.factor(frace),
         malform = as.factor(malform),
         mrace = as.factor(mrace))
```

There does not appear to be any missing values in the data. 

```{r}
sapply(birthweight_df, function(x) sum(is.na(x)))
```

In order to study the relationship between birthweight and several factors, we will fit a model with all variables and perform a stepwise regression to find an appropriate model.

#### Fitting a Model
```{r}
model = lm(bwt ~., birthweight_df)

step.model =  stepAIC(model, direction = "both", 
                      trace = FALSE)

step.model %>% 
  broom::tidy() %>% 
  dplyr::select(term, estimate, p.value) %>% 
  knitr::kable(digits = 3)
```

Running the stepwise regression, we will be studying the relationship between birthweight and the following independent variables: babysex, baby's head circumference, baby's length, mother's weight at delivery, family income, gestational age, mother's height, number of live births prior to this pregnancy, mother's pre-pregnancy weight, and average number of cigarettes smoked per day.

#### Plotting Residuals
```{r}
birthweight_df %>% 
  add_residuals(step.model) %>% 
  add_predictions(step.model) %>%
  ggplot(aes(x = pred, y = resid)) + 
  geom_point() +
  geom_hline(yintercept = 0, color = "red") +
  ggtitle("Residuals vs. Fitted Values") +
  xlab("Fitted Values") + ylab("Residuals")
```

Looking at the residuals vs. fitted values plot, it appears that for higher values of birthweight, the residuals are evenly distributed around 0. However, for lower values of birthweight, the residuals tend to be distributed greater than 0.

#### Comparing Models

In order to find the model that best predicts birthweight, we will use cross validation to compare the stepwise model with two other models. One model will look at length at birth and gestational age (main effects) while the other model will look at head circumference, length, sex, and all interactions.

```{r, warning = FALSE}
cv_df =
  crossv_mc(birthweight_df, 100) %>% 
  mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble))

cv_df = 
  cv_df %>% 
  mutate(
    step.model  = map(train, ~lm(bwt ~ 
      babysex + bhead + blength + delwt + fincome + 
      gaweeks + mheight + mrace + parity + ppwt + smoken, data = .x)),
    main_effects = map(train, ~lm(bwt ~ blength + gaweeks, data = .x)),
    interaction = map(train, ~lm(bwt ~ bhead * blength * babysex, data = .x))) %>%
  mutate(
    rmse_step = map2_dbl(step.model, test, ~rmse(model = .x, data = .y)),
    rmse_main   = map2_dbl(main_effects, test, ~rmse(model = .x, data = .y)),
    rmse_interaction = map2_dbl(interaction, test, ~rmse(model = .x, data = .y)))

cv_df %>% 
  dplyr::select(starts_with("rmse")) %>% 
  pivot_longer(
    everything(),
    names_to = "model", 
    values_to = "rmse",
    names_prefix = "rmse_") %>% 
  mutate(model = fct_inorder(model)) %>% 
  ggplot(aes(x = model, y = rmse)) + geom_violin()
```

According to the cross validation, it is apparent that the step-wise model has the lowest RMSE value while the main effects model has the highest RMSE. This means that it may be most appropriate to fit the step-wise model to predict birthweight. 


